---
title: "Untitled"
author: "RN7"
date: "May 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(spData)
library(sf)
library(tmap)
library(grid)

# best projection for inset? 
# equal area proj

# reproject us_states (from spData)

st_crs(us_states) # EPSG 4269

us_states2163 <- us_states %>% st_transform(crs = 2163)

st_crs(us_states2163) # EPSG 2163


```

Calculate scale relations between the mainland + hawaii and alaska
-> necessary to calculate the areas of the bounding box of each object with `st_area()`
--> then, calculate ratios

```{r}
us_states_area <- us_states2163 %>% st_bbox() %>% st_as_sfc() %>% st_area()
hawaii_area <- hawaii %>% st_bbox() %>% st_as_sfc() %>% st_area()
alaska_area <- alaska %>% st_bbox() %>% st_as_sfc() %>% st_area()

us_states_hawaii_ratio <- as.numeric(hawaii_area / us_states_area)
us_states_alaska_ratio <- as.numeric(alaska_area / us_states_area)


```

create individual maps for each of the objects: us-states, alaska, hawaii
-> use `tm_layout()` to remove map frames and backgrounds

```{r}
us_states_map <- tm_shape(us_states2163) + 
  tm_polygons() +
  tm_layout(frame = FALSE)

hawaii_map <- tm_shape(hawaii) +
  tm_polygons() +
  tm_layout(title = "Hawaii", 
            frame = FALSE,
            bg.color = NA,
            title.position = c("left", "BOTTOM"))

alaska_map <- tm_shape(alaska) +
  tm_polygons() +
  tm_layout(title = "Alaska", 
            frame = FALSE,
            bg.color = NA,
            title.position = c("left", "TOP"))


```

# Arrange map ----

use `grid` pkg to arrange map 

The code below creates a new ‘page’ and specifies its layout with two rows - one smaller (for Alaska) and one larger (for the contiguous 48 states). 

Each inset map is printed in its corresponding viewport, with Hawaii is printed in the bottom of the second row. 

Note the use of area rations to make the size of Alaska and Hawaii roughly comparable with that of the main land. The final step draws dashed lines `(gp = gpar(lty = 2))` to separate mainland from Alaska and Hawaii.


```{r}
grid.newpage()

pushViewport(viewport(layout = grid.layout(nrow = 2, ncol = 1, 
                                           heights = unit(c(us_states_alaska_ratio, 1), "null"))))

print(alaska_map, vp = viewport(layout.pos.row = 1))
print(us_states_map, vp = viewport(layout.pos.row = 2))
print(hawaii_map, vp = viewport(x = 0.25, y = 0.1, 
                                height = us_states_hawaii_ratio))

grid.lines(x = c(0, 1), y = c(0.64, 0.64), gp = gpar(lty = 2))
grid.lines(x = c(0, 0.5), y = c(0.32, 0), gp = gpar(lty = 2))



```


























