---
title: "Untitled"
author: "RN7"
date: "May 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

https://www.azavea.com/blog/2017/08/30/spatial-analysis-pipelines-in-r-with-simple-features/

FIveThrirtyEight >>> COngress voting Trump
+ 
shapefile of legislative districts from US Census Bureau

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(magrittr)

cd <- st_read("../congressional_districts.shp", stringsAsFactors = FALSE)

head(cd)

plot(cd)

```


load voting dataset

```{r}
cts <- read.csv("../congressional_trump_scores.csv", 
                stringsAsFactors = FALSE)[, -1] %>% 
  mutate(district = as.character(district))

glimpse(cts)



```

summary statistics already provided by FiveThirtyEight. 
->>> represent each Representative track record of voting relative to Trump's position on issue 
->>> "Trump Score": proportion of time each legislator voted in line with the president.

JOIN voting data with district polygons.

```{r}
class(cd) # sf + data.frame
class(cts) # data.frame

dat <- cd %>% left_join(cts)

class(dat) # sf + data.frame! 

```

map theme

```{r}
mapTheme <- function() {
  theme_void() + 
    theme(
      text = element_text(size = 7),
      plot.title = element_text(size = 11, color = "#1c5074", 
                                hjust = 0, vjust = 2, face = "bold"), 
      plot.subtitle = element_text(size = 8, color = "#3474A2", 
                                   hjust = 0, vjust = 0),
      axis.ticks = element_blank(), 
      legend.direction = "vertical", 
      legend.position = "right",
      plot.margin = margin(1, 1, 1, 1, 'cm'),
      legend.key.height = unit(1, "cm"), 
      legend.key.width = unit(0.2, "cm")
    ) 
}

```

first plot!

```{r}
dat %>% 
  ggplot() +
  geom_sf(aes(fill = trump_score), color = NA) +
  coord_sf(crs = st_crs(102003)) +
  scale_fill_gradient2("Trump Score \n", 
                       low = "#0099ff", mid = "#ffffff", high = "#ff6666",
                       midpoint = 0.5) +
  labs(
    title = 'Where have U.S. Representatives voted with and against President Trump?',
    subtitle = "Mapping FiveThirtyEight's 'Trump Score' of House of Representative's voting records",
    caption = "Source: Azavea, Data: FiveThirtyEight"
  ) +
  mapTheme()
```

EXTREME ends of spectrum == partisanship of congress!

Zoom in on Upper MidWest states (Minnesota, Wisconsin, Iowa)!


```{r}

upper_mw <- dat %>% 
  filter(state %in% c("MN", "IA", "WI")) %>% 
  st_transform(2289)

glimpse(upper_mw)

upper_mw_coords <- upper_mw %>% 
  st_centroid() %>%   # find centroids
  st_coordinates()    # extract coordinates as matrix

# find centroids for placement of labels!

upper_mw$long <- upper_mw_coords[, 1]
upper_mw$lat <- upper_mw_coords[, 2]

library(ggrepel)

upper_mw %>% 
  ggplot() +
  geom_sf(aes(fill = trump_score), color = "white") +
  coord_sf(datum = NA) +
  scale_fill_gradient2("Trump Score \n", 
                       low = "#0099ff", mid = "#ffffff", high = "#ff6666",
                       midpoint = 0.5) +
  geom_label(aes(long, lat, label = name, color = party),  
             alpha = 0.75, size = 2, inherit.aes = FALSE) +
  scale_color_manual("Political Party", 
                     values = c("Blue", "Red"),
                     guide = guide_legend(
                       override.aes = list(
                         shape = c(1, 1)))) +
  labs(
    title = "Congressional support for President Trump's policies in the Upper Midwest",
    subtitle = "Mapping FiveThirtyEight's 'Trump Score' of House of Representative's voting records",
    caption = "Source: Azavea, Data: FiveThirtyEight"
  ) +
  mapTheme() +
  guides(label = guide_legend(override.aes = list(shape = c(14, 1))))
  


```

- use `geom_text_repel()` for overlapping text lables 
- fix the goddamn legend for political party !!!!!

Aggregate original map >>> state-level! 

```{r}
by_state <- dat %>% 
  group_by(state) %>% 
  summarize(avg_trump_score = mean(na.omit(trump_score)),
            districts = n_distinct(district))

head(by_state)

```


plot ! 

```{r}
by_state %>% 
  ggplot() +
  geom_sf(aes(fill = avg_trump_score), color = "white") +
  scale_fill_gradient2("Average \nTrump Score \nby State \n",
                       low = '#0099ff', 
                       mid = '#ffffff', 
                       high = '#ff6666', 
                       midpoint = 0.5) +
  coord_sf(crs = st_crs(102003), datum = NA) +
  labs(
    title = "Which states' congressional delegations have supported President Trump's policies?",
    subtitle = "Mapping FiveThirtyEight's 'Trump Score' of House of Representative's voting records",
    caption = "Source: Azavea, Data: FiveThirtyEight"
  ) +
  mapTheme()





```




























