---
title: "Untitled"
author: "RN7"
date: "January 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2016 Presidential Election poll averages


http://www.thecrosstab.com/datascience/r-1/


```{r}
library(dplyr)
library(readr)

polls_2016 <- read_tsv(url("http://elections.huffingtonpost.com/pollster/api/v2/questions/16-US-Pres-GE%20TrumpvClinton/poll-responses-clean.tsv"))

glimpse(polls_2016)


```

1905 polls from the election

subset for entire electorate

```{r}
polls_2016 <- polls_2016 %>% 
  filter(sample_subpopulation %in% c("Adults", "Likely Voters", "Registered Voters"))

# 816 polls

unique(polls_2016$sample_subpopulation)

```

mutate date to proper format with `lubridate`

```{r}
library(lubridate)

glimpse(polls_2016)
# as date var

polls_2016 <- polls_2016 %>% 
  mutate(end_date = ymd(end_date))


```

expand data t ocover entire campaign duration

>>> moving average look at apst 2 weeks of poll results, even if not exist >> set to NA

```{r}
polls_2016 <- polls_2016 %>% 
  right_join(data.frame(
    end_date = seq.Date(min(polls_2016$end_date), max(polls_2016$end_date), by = "days")))

glimpse(polls_2016)

```


average the polls: rolling average of 14 polls over the campaign cycle over previous 2 weeks.


first: summarize for each day where multiple olls >>> average of polls 

```{r}
polls_2016 <- polls_2016 %>% 
  group_by(end_date) %>% 
  summarize(Clinton = mean(Clinton), 
            Trump = mean(Trump))
```


second: average polls with `rollapply()` from `zoo` package. 

variable to average >>> width of units over which to average (# of days), function to apply to specified units (# of days), how often to average, NAs?, fill NAs?, window function.

```{r}
library(zoo)

rolling_average <- polls_2016 %>% 
  mutate(Clinton.Margin = Clinton - Trump, 
         Clinton.Avg = rollapply(Clinton.Margin, width = 14, 
                                 FUN = function(x){mean(x, na.rm = TRUE)}, 
                                 by = 1, partial = TRUE, fill = NA, align = "right"))

tail(rolling_average)

```


ggplot

```{r}
library(ggplot2)

rolling_average %>% 
  ggplot() +
  geom_line(aes(end_date, Clinton.Avg), col = "blue") +
  geom_point(aes(end_date, Clinton.Margin))




```




# Use DT

```{r}

library(data.table)
pollsDT <- fread("http://elections.huffingtonpost.com/pollster/api/v2/questions/16-US-Pres-GE%20TrumpvClinton/poll-responses-clean.tsv")

pollsDT <- pollsDT[sample_subpopulation %in% c("Adults", "Likely Voters", "Registered Voters"), ]
pollsDT[, end_date := as.IDate(end_date)]
pollsDT <- pollsDT[data.table(end_date = seq(min(pollsDT[, end_date]),
                                  max(pollsDT[, end_date]), by = "days")), on = "end_date"]

# avg. polls

library(zoo)

pollsDT <- pollsDT[, .(Clinton = mean(Clinton), Trump = mean(Trump)), by = end_date]
glimpse(pollsDT)

pollsDT[, Clinton.Margin := Clinton - Trump]
glimpse(pollsDT)

pollsDT[, Clinton.Avg := rollapply(Clinton.Margin, width = 14, 
                                   FUN = function(x){mean(x, na.rm = TRUE)}, 
                                   by = 1, partial = TRUE, fill = NA, align = "right")]

glimpse(pollsDT)

# plot
library(ggplot2)

pollsDT %>% 
  ggplot() +
  geom_line(aes(end_date, Clinton.Avg), col = "blue") +
  geom_point(aes(end_date, Clinton.Margin))


```










