---
title: "Untitled"
author: "RN7"
date: "June 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

http://felixluginbuhl.com/mapping/



```{r}
library(tidyverse)
library(rvest)
library(readxl)


url_2016 <- "http://fundforpeace.org/fsi/wp-content/uploads/data/fsi-2016.xlsx"

# whatever it's not working so just download manually
index_2016 <- read_excel("data/fsi-2016.xlsx")

index_2016 %>% select(Country, Year, Rank, Total)


```

Create color palette

```{r}
library(RColorBrewer)

display.brewer.pal(10, "RdYlGn")

# break down continuous var into 10 discrete classes

breaks_map = c(0, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120)
labels_map = c("0-30", "30-40", "40-50", "50-60", "60-70", 
               "70-80", "80-90", "90-100", "100-110", "110-120")


index_2016 <- index_2016 %>% 
  mutate(Classes = cut(Total, 
                       breaks = breaks_map,
                       labels = labels_map))

```

Check countries have same names between Index and ggplot2 map data!

```{r}
world <- ggplot2::map_data("world") %>% 
  filter(region != "Antarctica")

# LOOK AT DIFFERENCES: setdiff()
dplyr::setdiff(index_2016$Country, world$region)

```

The following are different in index_2016 (LHS) from ggplot2::world (RHS):

[1] "Congo Democratic Republic" "Guinea Bissau"             "Cote d'Ivoire"            
 [4] "Congo Republic"            "Kyrgyz Republic"           "Israel and West Bank"     
 [7] "Brunei Darussalam"         "Trinidad and Tobago"       "Antigua and Barbuda"      
[10] "Slovak Republic"           "United States"             "United Kingdom"   

NEED to change index_2016 ^names to match the ones in RHS ggplot2::world

Use dplyr::recode() >>> similar to switch() but vectorized. For logical == if_else(), for more complicated critiera == case_when()

```{r}

index_2016 <- index_2016 %>%
  mutate(Country = recode(Country,
    "United States" = "USA",
    "United Kingdom" = "UK",
    "Slovak Republic" = "Slovakia",
    "Israel and West Bank" = "Israel",
    "Kyrgyz Republic" = "Kyrgyzstan",
    "Congo Democratic Republic" = "Democratic Republic of the Congo",
    "Congo Republic" = "Republic of Congo",
    "Antigua and Barbuda" = "Antigua",
    "Trinidad and Tobago" = "Trinidad",
    "Guinea Bissau" = "Guinea-Bissau",
    "Cote d'Ivoire" = "Ivory Coast",
    "Brunei Darussalam" = "Brunei"))

```

Create map!

```{r}
library(ggthemes)
library(ggalt)

ggplot() +
  geom_map(data = world, map = world, 
           aes(x = long, y = lat, map_id = region), 
           col = "lightblue", fill = "lightgrey", size = 0.1) +
  geom_map(data = index_2016, map = world, aes(map_id = Country, fill = Classes)) +
  coord_proj("+proj=wintri") + # Winkel Trimple projection
  scale_fill_brewer(NULL, palette = "RdYlGn", direction = -1) +
  theme_map() +
  theme(plot.title = element_text(face = "bold")) +
  labs(title = "Fragile States Index",
       subtitle = "2016 score (120=maximum)",
       caption = "Source: Fund for Peace, The Economist")


```

- more detailed with sf
- use countrycode pkg to replace names quickly!

```{r}
library(rnaturalearth)
library(countrycode)
library(sf)
library(hrbrthemes)

earth <- ne_countries(scale = 50, returnclass = "sf") %>% 
  select(admin, geometry, iso_a2)

glimpse(earth)

index_2016 <- index_2016 %>% 
  mutate(iso2 = countrycode(sourcevar = Country, origin = "country.name", destination = "iso2c"))

glimpse(index_2016)

data_map <- earth %>% 
  inner_join(index_2016, by = c("iso_a2" = "iso2")) %>% 
  select(Country, Total, Classes, geometry) %>% 
  sf::st_as_sf()

glimpse(data_map)

```

DRAW

```{r}
library(hrbrthemes)

data_map %>% 
  ggplot() +
  geom_sf(aes(fill = Classes), size = 0.1, color = "white") +
  coord_sf(crs = "+proj=moll") +        # Mollweide projection
  scale_fill_brewer(NULL,
                    palette = "RdYlGn", 
                    direction = -1,
                    guide = guide_legend(
                      title = "More to less stable",
                      title.hjust = 0.5, 
                      direction = "horizontal", 
                      nrow = 1, 
                      keyheight = unit(3, units = "mm"),
                      keywidth = unit(3, units = "mm"),
                      label.position = "right",
                      title.position = "top"
                    )) +
  theme_ipsum() +
  theme(legend.direction = "horizontal",
        legend.position = "bottom") +
  labs(title = "Fragile States Index",
       subtitle = "2016 score (120=maximum)",
       caption = "Félix Luginbühl (@lgnbhl)\nSource: Fund for Peace, The Economist")


```


LEAFLET 


```{r}
library(leaflet)

bins <- breaks_map 
pal <- colorBin("RdYlGn", domain = data_map$Total, bins = bins, reverse = TRUE)
labels <- sprintf("<strong>%s</strong><br/>%g Index score <sup></sup>",
                  data_map$Country, data_map$Total) %>% 
  lapply(htmltools::HTML) # turn text into HTML

leaflet_map <- data_map %>% 
  leaflet() %>% addTiles() %>% 
  addPolygons(
    fillColor = ~pal(data_map$Total), fillOpacity = 0.7,
    color = "white", weight = 1, opacity = 1, 
    highlight = highlightOptions(weight = 3),
    label = labels, 
    labelOptions = labelOptions(
      style = list("font-weight" = "normal"),
      textsize = "15px", direction = "auto")) %>% 
  addLegend(
    pal = pal, values = ~Classes, opacity = 0.9, 
    title = "Fragile Index", position = "bottomleft") %>% 
  addEasyButton(easyButton(icon = "fa-globe", title = "Zoom to World Level",
                           onClick = JS("function(btn, map){ map.setZoom(1);}")))





```




## DO SIMILAR FOR GPI DATA???
















