---
title: "Untitled"
author: "RN7"
date: "June 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(ggsoccer)
library(ggimage)
library(extrafont)
# loadfonts()

#                              2   1  
pass_data <- data.frame(x = c( 84, 82),
                        y = c(  6, 32),
                        x2 = c(77, 84),
                        y2 = c(13, 8))

#                            corner kick
curve_data <- data.frame(x = c(100),
                         y = c(0),
                         x2 = c(94),
                         y2 = c(35))
# golovin cross
cross_data <- data.frame(x = 76,
                         y = 19,
                         x2 = 94,
                         y2 = 60)

# Saudi failed clearance 
clearance_data <- data.frame(x = c(94),
                             y = c(35),
                             x2 = c(82),
                             y2 = c(33.5))

# Gazinski header
goal_data <- data.frame(x = c(94),
                        y = c(60),
                        x2 = c(99.2),
                        y2 = c(47.5))

# soccer ball image
goal_img <- data.frame(x = 100,
                       y = 47) %>% 
  mutate(image = "https://d30y9cdsu7xlg0.cloudfront.net/png/43563-200.png")

# golovin movement
golovin_data <- data.frame(x = c(83),
                           y = c(24.25),
                           x2 = c(77),
                           y2 = c(21))

# zhirkov movement 
zhirkov_data <- data.frame(x = 98,
                           y = 2,
                           x2 = 88,
                           y2 = 6)

g <- ggplot(pass_data) +
  annotate_pitch() +
  geom_segment(aes(x = x, y = y, xend = x2, yend = y2),
               arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
  geom_segment(data = clearance_data,
               aes(x = x, y = y, xend = x2, yend = y2
                   ), linetype = "dashed", size = 0.85) +
  geom_segment(data = goal_data,
               aes(x = x, y = y, xend = x2, yend = y2
                   ), linetype = "dashed", color = "red", size = 0.85) +
  geom_segment(data = golovin_data,
               aes(x = x, y = y, xend = x2, yend = y2
                   ), linetype = "dashed", color = "darkgreen", size = 1.2) +
  geom_segment(data = zhirkov_data,
               aes(x = x, y = y, xend = x2, yend = y2
                   ), linetype = "dashed", color = "darkgreen", size = 1.25) +
  geom_curve(data = curve_data, 
             aes(x = x, y = y, xend = x2, yend = y2), 
             curvature = 0.25, 
             arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
  geom_curve(data = cross_data, 
             aes(x = x, y = y, xend = x2, yend = y2), 
             curvature = 0.25, 
             arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
  geom_image(data = goal_img,
             aes(x = x, y = y,
                 image = image), size = 0.035) +
  theme_pitch() + 
  theme(text = element_text(family = "Trebuchet MS")) +
  coord_flip(xlim = c(49, 101),
             ylim = c(-1, 101)) +
  ggtitle(label = "Russia (5) vs. (0) Saudi Arabia", 
          subtitle = "First goal, Yuri Gazinsky (12th Minute)") +
  labs(caption = "By Ryo Nakagawara (@R_by_Ryo)") +
  geom_label(aes(x = 94, y = 60, label = "Gazinsky"), hjust = -0.1) +
  geom_label(aes(x = 83, y = 23, label = "Golovin"), hjust = -0.05) +
  geom_label(aes(x = 75, y = 11, label = "Golovin"), hjust = -0.1) +
  geom_label(aes(x = 98, y = 0, label = "Zhirkov"), vjust = -0.3) +
  geom_label(aes(x = 84, y = 6, label = "Zhirkov"), vjust = -0.3) +
  annotate("text", x = 69, y = 65, family = "Trebuchet MS",
           label = "After a poor corner kick clearance\n from Saudi Arabia, Golovin picks up the loose ball, \n exchanges a give-and-go pass with Zhirkov\n before finding Gazinsky with a beautiful cross!")

ggsave(g, filename = "gazinsky_goal.png")

```




## Cristiano Ronaldo hattrick


```{r}
library(tidyverse)
library(ggsoccer)
library(extrafont)
library(emoGG)
library(ggimage)
library(ggflags)
# loadfonts()
# Official WC 2018 Font: "Dusha"
# http://fifa2018wiki.com/fifa-2018-font-typeface-download-dusha-font-ttf/509/

emoji_search("soccer")  # "26bd"

goals_data <- data.frame(x = c(88, 80, 71),
                         y = c(50, 48, 54))

cr <- ggplot(goals_data) +
  annotate_pitch() +
  theme_pitch() +
  theme(text = element_text(family = "Dusha V5")) +
  coord_flip(xlim = c(55, 112),
             ylim = c(-1, 101)) +
  geom_curve(x = 88, y = 50, 
             xend = 100, yend = 54,     # Penaldo
             curvature = 0.3, 
             arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
  geom_segment(x = 80, y = 48, 
               xend = 97, yend = 48) +  # 2nd 
  geom_segment(x = 97, y = 48, 
               xend = 100, yend = 45.5,
               arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +        # degea fumble
  geom_curve(x = 71, y = 54, 
             xend = 100, yend = 54,     # FREEKICK
             curvature = 0.3, 
             arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
  annotate(geom = "text", family = "Dusha V5", hjust = c(0.5, 0.5, 0.5, 0, 0, 0),
           size = c(6.5, 4.5, 3, 3.5, 3.5, 3.5),
           label = c("Portugal           (3) vs. Spain           (3)",
                     "Cristiano's Hattrick (4', 44', 88')",
                     "by Ryo Nakagawara (@R_by_Ryo)",
                     "1. Fouled by Nacho in the box,\nCristiano confidently strokes the ball\ninto the right corner from the spot.",
                     "2. Guedes lays it off to Cristiano whose\nstrong shot is uncharacteristically\nfumbled by De Gea into the net.",
                     "In the final minutes of the game,\nCristiano wins a freekick against Pique\nand curls it beautifully over the wall."),
           x = c(110, 105, 53, 76, 66, 66), 
           y = c(30, 20, 85, 5, 5, 55)) +
  ggimage::geom_flag(aes(image = "PT"),       # Portugal Flag
            x = 110, y = 19, size = 0.08) +
  ggimage::geom_flag(aes(image = "ES"),       # Spain Flag
            x = 110, y = 51.5, size = 0.08) +
  ggimage::geom_emoji(aes(x = 105, 
                          y = c(45, 50, 55)),
             image = "26bd", size = 0.035) +
  geom_point(aes(x = x, y = y), 
             shape = 21, size = 7, color = "black", fill = "white") +
  geom_text(aes(x = x, y = y, label = c(1, 2, 3)), family = "Dusha V5")

ggsave(cr, filename = "cr_hattrick.png")

```


"1. Fouled by Nacho in the box, Cristiano confidently strokes the ball into the right corner from the spot."

"2. Guedes lays it off to Cristiano who's strong shot is uncharacteristically fumbled by De Gea into the net."

"3. In the final minutes of the game, Cristiano wins a freekick against Pique and curls it beautifully over the wall."








## Shots on Target  // Off Target

```{r}

```



## Formation and line-ups



```{r}

```

















