---
title: "Untitled"
author: "RN7"
date: "May 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

https://www.jessesadler.com/post/gis-with-r-intro/


```{r}
# Load the packages
library(tidyverse)
library(sp)
library(sf)
library(rnaturalearth)

# Load the data
letters <- read_csv("../correspondence-data-1585.csv")
locations <- read_csv("../locations.csv")
```

create calculated dfs

```{r}
# letters per source
sources <- letters %>% 
  group_by(source) %>% 
  count() %>% 
  rename(place = source) %>% 
  add_column(type = "source") %>% 
  ungroup()

# letters per destination
destinations <- letters %>% 
  group_by(destination) %>% 
  count() %>% 
  rename(place = destination) %>% 
  add_column(type = "destination") %>% 
  ungroup()



```

join together

```{r}

letters_data <- rbind(sources, destinations) %>% 
  mutate(type = as_factor(type))

glimpse(letters_data)

```

join to locations by "place"

```{r}
geo_data <- letters_data %>% left_join(locations, by = "place")

glimpse(geo_data)

```

sp pkg

CRS == WGS84

```{r}
coords <- geo_data %>% select(lon, lat)

points_sp <- SpatialPoints(coords = coords, 
                           proj4string = CRS("+proj=longlat +datum=WGS84"))

points_sp
glimpse(points_sp)

```

points_sp NOT contain non-spatial data (from letters_data) >>>
create SpatialPointsDataFrame object >>> fill data slot with letters_data

```{r}
points_spdf <- SpatialPointsDataFrame(coords = coords, 
                                      data = letters_data, 
                                      proj4string = CRS("+proj=longlat +datum=WGS84"))

points_spdf

str(points_spdf, max.level = 2)

```

Access non-spatial data with `points_spdf@data` >>> inability use dplyr commands

```{r}
points_spdf[points_spdf@data$n > 10, ]
```

Grab map data

```{r}
# rnaturalearth

coast_sp <- ne_coastline(scale = "medium")
countries_sp <- ne_countries(scale = "medium")

str(coast_sp, max.level = 2)
str(countries_sp, max.level = 2)

```

# make sure points_spdf and the map sp dfs have SAME CRS.

```{r}
coast_sp@proj4string
# WGS84

proj4string(countries_sp)
# WGS84

proj4string(points_spdf)
3=# WGS84
```

use base plotting for sp objects

```{r}
palette(alpha(c("darkorchid", "darkorange"), 0.7))

par(mar = c(1, 1, 3, 1))

plot(points_spdf, 
     pch = 20, col = points_spdf$type,
     cex = sqrt(points_spdf$n) / 2 + 0.25)

box()

title(main = "Correspondence of D. vdMeulen, 1585")

```


```{r}
pointsize <- c(1, 50, 100)
par(mar = c(1, 1, 3, 1))

# Plot points
plot(points_spdf,
     pch = 20,
     col = points_spdf$type,
     cex = sqrt(points_spdf$n)/2 + 0.25)
# Plot coastlines background map
plot(coast_sp,
     col = "black",
     add = TRUE)
# Add a box around the plot
box()

# Legend for colors
legend("topright", legend = levels(points_spdf$type),
       pt.cex = 2,
       col = 1:2,
       pch = 15)

# legend for size of points
legend("right", legend = pointsize,
       pt.cex = (sqrt(pointsize)/2 + 0.25),
       col = "black",
       pch = 20,
       title = "Letters")

# Title for the map
title(main = "Correspondence of Daniel van der Meulen, 1585")
```


to link countries with map so far >>> change bounding box to match points_spdf with `bbox()`

```{r}
countries_sp@bbox <- bbox(points_spdf)


par(mar = c(1, 1, 3, 1))

# Plot countries map and color with grays
plot(countries_sp,
     col = gray(0.8),
     border = gray(0.7))
# Plot points
plot(points_spdf,
     pch = 20,
     col = points_spdf$type, 
     cex = sqrt(points_spdf$n)/2 + 0.25,
     add = TRUE)
# Add a box around the plot
box()

# Legend for colors
legend("topright",
       legend = levels(points_spdf$type),
       pt.cex = 2,
       col = 1:2,
       pch = 15)
# legend for size of points
legend("right",
       legend = pointsize,
       pt.cex = (sqrt(pointsize)/2 + 0.25),
       col = "black",
       pch = 20,
       title = "Letters")

# Title for the map
title(main = "Correspondence of Daniel van der Meulen, 1585")

```


## same workflow but with sf!

```{r}
points_sf <- st_as_sf(geo_data, 
                      coords = c("lon", "lat"),
                      crs = 4326)

class(points_sf)
points_sf

class(points_sf$geometry)

# geometry of the sf object
st_geometry(points_sf)

```

rnaturalearth >> create into sf with returnclass arg

```{r}
coast_sf <- ne_coastline(scale = "medium", returnclass = "sf")
countries_sf <- ne_countries(scale = "medium", returnclass = "sf")

head(coast_sf)

# subset with tidyverse verbs!

points_sf %>% filter(n > 10)

# manipulate sf with tidyverse verbs
countries_sf %>% 
  filter(continent == "South America") %>% 
  select(name) %>% 
  st_transform(crs = "+proj=moll +datum=WGS84") %>% 
  plot(key.pos = NULL, graticule = TRUE, main = "SA")

```

sf and ggplot2

If you make a ggplot2 plot with points_sf and coast_sf without modifications, you will get a wold map with some small points in western Europe. There are two ways to deal with this incongruence by either changing the bounding box of the background map before making the plot or within the plot itself. As far as I know, there is no equivalent method in the sf package to the command we used above to modify the bbox slot of the countries_sp object. However, it is possible to geographically subset an sf object by a bounding box of coordinates using the ms_clip() function from the rmapshaper package. The other option is to leave coast_sf and countries_sf as they are and adjust the geographic extent of the objects in the plot command itself with the coord_sf() function and the xlim and ylim arguments. Here, I will use the latter method.

There are some oddities in using geom_sf() compared to other geoms due to the character of geom_sf(). By its nature, geom_sf() is unlike other geoms in that it can be used to create points, lines, or polygons depending on the contents of the sf object. This variability results in the need to declare the geometrical object used in a show.legend argument for the legend to be properly formatted. Another difference that I have found from other geoms is that geom_sf() does not properly identify sf objects unless you explicitly identify them with data = sf_object. While these issues are something to be aware of, they do not substantively affect the the plots that we can create.


```{r}
ggplot() +
  geom_sf(data = coast_sf) +
  geom_sf(data = points_sf, 
          aes(color = type, size = n), 
          alpha = 0.7, 
          show.legend = "point") +  # specify legend format
  coord_sf(xlim = c(-1, 14), 
           ylim = c(44, 55))
```

On the other hand, it is possible to add information to the plot by labeling the points with geom_text_repel() from the ggrepel package, which will ensure that the labels for the points do not overlap. geom_text_repel() does not work with sf objects, but because we have made no geographic transformations to the data, we can use the data from the locations data frame to map the names of the locations to the longitude and latitude values.

```{r}
library(ggrepel)

ggplot() +
  geom_sf(data = coast_sf) +
  geom_sf(data = points_sf,
          aes(color = type, size = n), 
          alpha = 0.7, 
          show.legend = "point") +
  coord_sf(xlim = c(-1, 14), ylim = c(44, 55),
           datum = NA) + # remove graticules (the coord lines thingys)
  labs(title = "Correspondence of Daniel van der Meulen, 1585",
       size = "Letters",
       color = "Type",
       x = NULL,
       y = NULL) +
  guides(color = guide_legend(override.aes = list(size = 6))) +
  theme_minimal()


```

fill in land, use countries_sf for borders

```{r}
ggplot() +
  geom_sf(data = countries_sf,
          fill = gray(0.8), 
          color = gray(0.7)) +
  geom_sf(data = points_sf,
          aes(color = type, size = n), 
          alpha = 0.7, 
          show.legend = "point") +
  coord_sf(xlim = c(-1, 14), ylim = c(44, 55),
           datum = NA) + # remove graticules (the coord lines thingys)
  labs(title = "Correspondence of Daniel van der Meulen, 1585",
       size = "Letters",
       color = "Type",
       x = NULL,
       y = NULL) +
  guides(color = guide_legend(override.aes = list(size = 6))) +
  theme_bw()


```






















