---
title: "Untitled"
author: "RN7"
date: "June 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

http://uribo.hatenablog.com/entry/2017/10/12/083110

Labelling GoogleMaps + OpenStreetMap features.



Use: 
https://cyberjapandata.gsi.go.jp/xyz/experimental_nrpt/{z}/{x}/{y}.geojson

From GSI Japan

```{r}
latlon2tile <- function(lon, lat, z) {
    x = trunc((lon/180 + 1) * 2^z/2)
    y = trunc(((-log(tan((45 + lat/2) * pi/180)) + 
        pi) * 2^z/(2 * pi)))
    return(list(x = x, y = y, zoom = z))
}
```

use the function above to create coordinates for the area we want to map.

```{r}
(xyz <- latlon2tile(133.938, 34.6636, 15))
```

Note that the VectorTile data is anchored at zoom level 15.

Read in the .geojson files with the sf package.

```{r}
library(sf)

# 居住地名
df.nrpt <- read_sf(paste0("https://cyberjapandata.gsi.go.jp/xyz/experimental_nrpt/", 
    paste(xyz[3], xyz[1], xyz[2], sep = "/"), 
    ".geojson"))

# 公共施設
df.nffpt <- read_sf(paste0("https://cyberjapandata.gsi.go.jp/xyz/experimental_pfpt/", 
    paste(xyz[3], xyz[1], xyz[2], sep = "/"), 
    ".geojson"))

dplyr::glimpse(df.nrpt)


```

Download Okayama City shapefiles from estat.jp!

```{r}
library(dplyr)
library(purrr)

pref33 <- read_sf("../A002005212015DDSWC33102") %>% 
  rbind(read_sf("../A002005212015DDSWC33101")) %>% 
  st_transform(crs = 4326) %>% 
  select(MOJI)

glimpse(pref33)

```

Filter the shapefile data so we only use the areas we have in our VectorTiles data. 

Use `st_contains()` to find the specific points and polygons.

```{r}
which.row <- data_frame(contains = pref33$geometry %>% 
                          map(~st_contains(.x, df.nrpt$geometry)) %>% 
                          flatten() %>% 
                          map_lgl(~!identical(.,
                                              integer(0)))) %>% 
  mutate(row_num = row_number()) %>% 
  filter(contains == TRUE) %>% 
  magrittr::use_series(row_num) # only select row_num vector

glimpse(which.row)

pref33mod <- pref33 %>% slice(which.row)

glimpse(pref33mod)

```


PLOT

first, create centroid for labels

```{r}
df.nrpt %>% glimpse()

df.nrpt %<>% 
  select(name, kana) %>% 
  mutate(lon = map_dbl(geometry, ~st_centroid(.x)[[1]]), 
         lat = map_dbl(geometry, ~st_centroid(.x)[[2]]))

glimpse(df.nrpt)

df.nffpt %<>% 
  select(type, name = pfName) %>% 
  mutate(lon = map_dbl(geometry, ~st_centroid(.x)[[1]]), 
         lat = map_dbl(geometry, ~st_centroid(.x)[[2]]))

glimpse(df.nffpt)

```

PLOT

```{r}
library(ggplot2)
library(ggrepel)

ggplot() +
  geom_sf(data = pref33mod, fill = "white") +
  geom_text(data = df.nrpt, 
            aes(label = name, x = lon, y = lat, alpha = 0.5), 
            family = "PIAexGothic", size = 3) +
  geom_sf(data = df.nffpt, 
          aes(color = "red")) +
  geom_text_repel(data = df.nffpt, 
                  aes(label = name, x = lon, y = lat),
                  family = "IPAexGothic", size = 2) +
  theme(legend.position = "none") # OR use guides(alpha = FALSE, color = FALSE)


```


PLOT with Map Symbols!

```{r warning=FALSE}
library(emojifont)
load.fontawesome()

df.nffpt$label <- fontawesome(c("fa-dot-circle-o", "fa-plus-square",
                                "fa-times-circle", 
                    "fa-times-circle", "fa-university",
                    "fa-envelope", "fa-envelope"))

ggplot() +
  geom_sf(data = pref33mod, fill = "white") +
  geom_sf(data = df.nffpt, aes(color = "red")) +
  geom_label_repel(data = df.nffpt,
            aes(label = label, x = lon, y = lat), 
            family = "fontawesome-webfont") +
  geom_text(data = df.nrpt, aes(label = name, x = lon, y = lat,
                                alpha = 0.5), 
            family = "IPAexGothic", size = 3) + 
  geom_text_repel(data = df.nffpt, aes(label = name, x = lon, y = lat), 
             size = 2,
             family = "IPAexGothic") +
  guides(alpha = FALSE, color = FALSE)



```


























