---
title: "Untitled"
author: "RN7"
date: "February 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Len Kiefer: mortgage rate increases graphs 

http://lenkiefer.com/2018/02/05/comparing-recent-periods-of-mortgage-rate-increases/

```{r}
library(tidyverse)
library(readxl)
library(tidyquant)
library(tibbletime)
library(ggridges)
library(viridis)
library(cowplot)
library(ggbeeswarm)

```


Mortgage data from freddiemac.com: http://www.freddiemac.com/pmms/docs/historicalweeklydata.xls

```{r}
tickers <- c("MORTGAGE30US")

df <- tq_get(tickers, get = "economic.data", from = "2000-01-01") # get data in tibble format

df <- tq_get(tickers, get = "economic.data", from = "2012-01-01")

```

ID periods of rapid rate increases! Use algorithmic apporaches? (quantmod::findPeaks()), etc.

```{r}

df_1 <- df %>% 
  filter(date >= "2013-05-02" & date <= "2013-09-12") %>% 
  mutate(period = "Taper Tantrum 2013", 
         id = row_number())

glimpse(df_1)

df_2 <- df %>% 
  filter(date >= "2016-10-27" & date <= "2017-03-16") %>% 
  mutate(period = "US Election 2016", 
         id = row_number())

glimpse(df_2)

df_3 <- df %>% 
  filter(date >= "2017-09-14" & date <= "2018-03-31") %>% 
  mutate(period = "Recent period", 
         id = row_number())

glimpse(df_3)

df_note <- data.frame(date = as.Date(c("2013-05-02", "2016-10-27", "2017-09-14")),
                      period = c("Taper Tantrum 2013", "US Election 2016", "Recent period"),
                      end_date = as.Date(c("2013-09-12", "2017-03-16", "2018-03-31")))

glimpse(df_note)

df2 <- bind_rows(df_1, df_2, df_3)
glimpse(df2)

df2 <- df2 %>% mutate(periodf = factor(period, 
                                       levels = c("Taper Tantrum 2013", 
                                                  "US Election 2016", "Recent period")))

df_note_1 <- df_note %>% left_join(df2 %>% select(date, price, id, periodf), by = "date")
glimpse(df_note_1)

df_note_1 <- df_note_1 %>% left_join(df2 %>% select(date, price, id), by = c("end_date" = "date"))

```


```{r}
# graph
g1 <- df2 %>% 
  ggplot(aes(id, price, color = periodf, fill = periodf)) +
  geom_ribbon(aes(ymin = 3.3, ymax = price), color = NA, alpha = 0.2) +
  geom_line(size = 1.1) +
  scale_fill_viridis(option = "D", discrete = TRUE, end = 0.85) +
  scale_color_viridis(option = "D", discrete = TRUE, end = 0.85) +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(1, 23)) +
  scale_y_continuous(limits = c(3.3, 4.7), breaks = seq(3.3, 4.7, 0.2), sec.axis = dup_axis()) +
    labs(x = "weeks since rates started rising",y="30-year fixed mortgage rates (%)",
       title = "Recent periods of mortgage rate increases",
       subtitle = "U.S. weekly average 30-year fixed mortgage rates (%)",
       caption = "@lenkiefer Source: Freddie Mac Primary Mortgage Market Survey through February 1, 2018") +
  ggrepel::geom_label_repel(data = df_note_1, inherit.aes = FALSE, fontface = "bold", size = 3,
                            aes(x = id.y, color = periodf, label = period, y = price.y)) +
  ggrepel::geom_label_repel(data = df_note_1, inherit.aes = FALSE, fontface = "bold", size = 3,
                            aes(x = id.x, color = periodf, 
                                label = paste0("Began: ", as.character(date, 
                                                                       format = "%b %d, %Y")), 
                                y = price.x)) +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "none", 
        plot.title = element_text(face = "bold", size = 18),
        plot.subtitle = element_text(face = "italic", size = 14),
        strip.text = element_text(face = "bold", size = 20)) 

g1

```



recent period - time series plot

```{r}
g2 <- df %>% 
  ggplot(aes(date, price)) +
  geom_line(color = "darkgrey", inherit.aes = FALSE, 
            aes(date, price)) +
  theme_minimal() +
  scale_y_continuous(limits = c(3.3, 4.7), breaks = seq(3.3, 4.7, 0.2), 
                     sec.axis = dup_axis()) +
  scale_fill_viridis(option = "D", discrete = TRUE, end = 0.85) +
  scale_color_viridis(option = "D", discrete = TRUE, end = 0.85) +
  labs(x = "", y = "30-year fixed mortgage rates (%)", 
       title = "Recent periods of mortgage rate increases", 
       subtitle = "US weekly avg. 30-year fixed mortgage rates (%)", 
       caption = "@lenkiefer, Source: Freddie Mac Primary Mortgage Market Survey through Feb. 1, 2018") +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "none", 
        plot.title = element_text(face = "bold", size = 18), 
        plot.subtitle = element_text(face = "bold", size = 14), 
        strip.text = element_text(face = "bold", size = 14))


g2

```

"eyeball" econometrics >>> 3 distinct periods of rate increases!

```{r}
g3 <- 
  df2 %>% 
  ggplot(aes(date, price, color = periodf, fill = periodf)) +
  geom_line(data = df %>% mutate(id = row_number()), color = "darkgray", inherit.aes = FALSE,
            aes(date, price)) +
  geom_ribbon(aes(ymin = 3.3, ymax = price), color = NA, alpha = 0.2) + 
  geom_line(size = 1.1) +
  scale_fill_viridis(option = "D", discrete = TRUE, end = 0.85) +
  scale_color_viridis(option = "D", discrete = TRUE, end = 0.85) +
  theme_minimal() +
  geom_label(data = df_note_1, inherit.aes = FALSE, fontface = "bold", size = 3,
             aes(end_date, color = periodf, label = period, y = price.y), 
             hjust = 1, nudge_y = 0.05) +
  scale_y_continuous(limits = c(3.3, 4.7), breaks = seq(3.3, 4.7, 0.2), sec.axis = dup_axis()) +
  labs(x = "", y = "30-year fixed mortgage rates (%)",
       title = "Recent periods of mortgage rate increases",
       subtitle = "U.S. weekly average 30-year fixed mortgage rates (%)",
       caption = "@lenkiefer Source: Freddie Mac Primary Mortgage Market Survey through February 1, 2018") +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 18), 
        plot.subtitle = element_text(face = "bold", size = 14), 
        strip.text = element_text(face = "bold", size = 14))
  
g3

```

ow use small multiples to take closer look at above! 

```{r}
g4 <- df2 %>% 
  ggplot(aes(x=id,y=price, color=periodf,fill=periodf))+
  geom_ribbon(aes(ymin=3.3,ymax=price),color=NA,alpha=0.2)+
  geom_line(size=1.1)+
  scale_fill_viridis(option="D",discrete=T,end=0.85)+
  scale_color_viridis(option="D",discrete=T,end=0.85)+
  theme_minimal()+
  scale_x_continuous(expand=c(0,0),limits=c(1,23))+
  ggrepel::geom_label_repel(data=df_note_1, inherit.aes=F, fontface="bold",size=3,
                            aes(x=id.y,color=periodf, label=paste0(as.character(end_date,format="%b %d,%Y"),": ",price.y),y=price.y))+
  ggrepel::geom_label_repel(data=df_note_1, inherit.aes=F, fontface="bold",size=3,
                            aes(x=id.x,color=periodf, label=paste0(as.character(date,format="%b %d,%Y"),": ",price.x),y=price.x))+
  scale_y_continuous(limits=c(3.3,4.7),breaks=seq(3.3,4.7,.2),sec.axis=dup_axis())+
  labs(x="weeks since rates started rising",y="30-year fixed mortgage rates (%)",
       title="Recent periods of mortgage rate increases",
       subtitle="U.S. weekly average 30-year fixed mortgage rates (%)",
       caption="@lenkiefer Source: Freddie Mac Primary Mortgage Market Survey through February 1, 2018")+
  theme(plot.caption=element_text(hjust=0),legend.position="none",
        plot.title=element_text(face="bold",size=18),
        plot.subtitle=element_text(face="italic",size=14),
        strip.text=element_text(face="bold",size=14)) +
  facet_wrap(~periodf)

g4

```

Arrange into single panel:

```{r}
cowplot::plot_grid(g1 + labs(caption = ""), # erase caption from g1
                   g4 + theme(
                     plot.title = element_blank(), 
                     plot.subtitle = element_blank()),     # erase title/subtitle from g4
                   ncol = 1)                           # specify 1 column
```


Cividis color scheme for colorblindess. `Option = "E"`

```{r}
cowplot::plot_grid(g1+labs(caption="")+ # remove caption   
                     scale_fill_viridis(option="E",discrete=T,end=0.85)+ # use cividis color map
                     scale_color_viridis(option="E",discrete=T,end=0.85),
                   g4+theme(plot.title=element_blank(), plot.subtitle=element_blank())+
                     scale_fill_viridis(option="E",discrete=T,end=0.85)+ # use cividis color map
                     scale_color_viridis(option="E",discrete=T,end=0.85),
                   ncol=1)
```



