---
title: "Untitled"
author: "RN7"
date: "July 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## http://lenkiefer.com/2018/07/26/getting-animated-about-new-home-sales/


load packages

```{r}
suppressPackageStartupMessages({
  library(data.table)
  library(tidyverse)
  library(viridis)
  library(ggridges)
  library(extrafont)
  library(lubridate)
  library(forcats)
  # for animations
  library(tweenr)
  library(gifski)}
  )
```



load data

```{r}
dt <- fread("data/RESSALES-mf.csv", skip = 719) # skip to row 719

glimpse(dt)
```

- per_idx is a monthly index starting in January 1963
Add dates

```{r}
df.dates <- data.frame(
  per_idx = 1:666,
  date = seq.Date(from = as.Date("1963-01-01"), by = "1 month", length.out = 666))

df <- dt %>% left_join(df.dates, by = "per_idx")

glimpse(df)
```

plot trend for new home sales
- include band for uncertainty around estimate
- Census provide relative standard error >>> % of point estimate
- normal approximation to construct band around monthly estimate
- USA: geo_idx = 1
- total (all houses): dt_idx = 1 
- seasonally adjusted annual rate: cat_idx = 2

```{r}
df2 <- df %>% 
  filter(cat_idx == 2, dt_idx == 1, geo_idx == 1)

# errors
df.err <- df %>% 
  filter(et_idx == 1, cat_idx == 2, geo_idx == 1)

```

- relative standard error: et_idx
- convert to CI by `qnorm()` for compute percentiles (assume norm distr centered at estimate with SE == estimates * et_idx / 100)

```{r}
df3 <- data.frame(
  date = df.err$date,
  err = df.err$val
)

df3 <- df3 %>% left_join(df2, by = "date")

```

## plot

```{r}
df3 %>% 
  filter(year(date) > 2014) %>% 
  ggplot(aes(x = date, y = val, fill = "5th to 25th (75th to 95th)")) +
  scale_fill_viridis(option = "C", discrete = T, direction = 1, 
                     begin = 0.85, end = 0.9,
                     name = "Confidence Interval Around Estimates") +
  geom_ribbon(aes( # 5-25 percentile
    ymin = qnorm(0.05, val, val * err / 100), 
    ymax = qnorm(0.25, val, val * err / 100)),
    alpha = 0.2) +
  geom_ribbon(aes( # 25-75 percentile
    ymin = qnorm(.25, val, val * err / 100), 
    ymax = qnorm(.75, val, val * err / 100), 
    fill = "25th to 75th pctile"), 
    alpha = 0.5) +
  geom_ribbon(aes( # 75-95 percentile
    ymin = qnorm(.75, val, val * err / 100), 
    ymax = qnorm(.95, val, val * err / 100)), 
    alpha = 0.2) +
  geom_line(size = 1.1, color = "#27408b") +
  geom_hline(data = . %>% tail(1), # add line for CURRENT
             aes(yintercept = val), linetype = 2) + 
  scale_y_continuous(sec.axis = dup_axis()) + # add dupe axis
  theme_ridges(font_family = "Roboto") +
  theme(legend.position = "top", plot.title = element_text(size = 20)) +
  labs(x = "", y = "",
       subtitle = "New Home Sales (1000s, SAAR)",
       title = "U.S. New Home Sales",
       caption = "@lenkiefer Source: U.S. Census Bureau and Department of Housing and Urban Development\nConfidence interval based on normal distribution given relative standard error.\nDotted line at June 2018 estimate")
```


WIDE CI

## Regional trends

```{r}
df2.r <- filter(df, 
                cat_idx==2,   # NSA home sales
                dt_idx==1,    # All houses
                ! geo_idx ==1 # not region =1 (excludes US)
                )

df.err.r <- filter(df, et_idx==1, cat_idx==2,! geo_idx ==1)

df3.r <- data.frame(date=df.err.r$date,err=df.err.r$val,geo_idx=df.err.r$geo_idx)

df3.r <- left_join(df3.r, df2.r, by=c("date","geo_idx")) 


df3.r <- 
  df3.r %>%
  mutate(mname=fct_reorder(factor(as.character(date, format="%b")),-month(date)),
         year=year(date),
         region=case_when(
           geo_idx==1 ~ "US",
           geo_idx==2 ~ "Northeast",
           geo_idx==3 ~ "Midwest",
           geo_idx==4 ~ "South",
           geo_idx==5 ~ "West"
         ))

```

### plot

```{r}
ggplot(data=filter(df3.r,year(date)>2014), 
       aes(x=date, y=val, fill="5th to 25th (75th to 95th)"))+
  scale_fill_viridis(option="C", discrete=T, direction=1, 
                     begin=0.85,end=0.9, name="Confidence Interval Around Estimates")+
  geom_ribbon(aes(
    ymin=qnorm(.05,val,val*err/100), 
    ymax=qnorm(.25,val,val*err/100)), alpha=0.25)+
  geom_ribbon(aes(
    ymin=qnorm(.25,val,val*err/100), 
    ymax=qnorm(.75,val,val*err/100), fill="25th to 75th pctile"), alpha=0.75)+
  geom_ribbon(aes(
    ymin=qnorm(.75,val,val*err/100), 
    ymax=qnorm(.95,val,val*err/100)), alpha=0.25)+
  facet_wrap(~paste0(region," Region"),scales="free")+
  geom_line(size=1.05,color="#27408b")+
  geom_hline(data= .%>% filter(date==max(df3$date)), 
             aes(yintercept=val),linetype=2)+
  scale_y_continuous(sec.axis=dup_axis())+
  theme_ridges(font_family="Roboto")+
  theme(legend.position="top", plot.title=element_text(size=20))+
  labs(x="", y="",
       subtitle="New Home Sales (1000s, SAAR)\nNote different scale for each panel!",
       title="U.S. Regional New Home Sales",
       caption="@lenkiefer Source: U.S. Census Bureau and Department of Housing and Urban Development\nConfidence interval based on normal distribution given relative standard error.\nDotted line at June 2018 estimate")
```


## Year-to-Date totals + compare across years


```{r}
df.nsa <- df %>% 
  filter(cat_idx == 1, dt_idx == 1, geo_idx == 1) %>% 
  select(date, geo_idx, val) %>% 
  mutate(mname = fct_reorder(factor(as.character(date, format = "%b")), -month(date)),
         year = year(date)) %>% 
  group_by(year) %>% 
  mutate(vsum = cumsum(val)) %>% # cumulative values of # of new home sales
  ungroup()
```

### plot

```{r}
df.nsa %>% 
  filter(year <= 21999, year > 1989, month(date) <= 6) %>% 
  ggplot(aes(x = factor(year), y = val, fill = mname, label = mname)) +
  geom_col(alpha = 0.7) +
  coord_flip() +
  geom_text(aes(y = vsum), hjust = 1, color = "white") +
  scale_fill_viridis_d(option = "C", end = 0.9, name = "Month") +
  theme_ridges(font_family = "Roboto Condensed", font_size = 14) +
  theme(legend.position = "none") +
  geom_hline(data = . %>% tail(1), 
             aes(yintercept = vsum), linetype = 2) +
  labs(x="year",y="New Home Sales (1000s NSA)",
       title="U.S. New Home Sales",
       subtitle="Year-to-date total through June",
       caption="@lenkiefer Source: U.S. Census Bureau and Department of Housing and Urban Development\ndotted line at 2018 value") + 
  guides(colour = guide_legend(reverse=FALSE))
```





## Animated


```{r}
library(gganimate)

df.nsa %>% 
  filter(year <= 21999, year > 1989, month(date) <= 6) %>% 
  ggplot(aes(x = factor(year), y = val, fill = mname, label = mname)) +
  geom_col(alpha = 0.7) +
  coord_flip() +
  geom_text(aes(y = vsum), hjust = 1, color = "white") +
  scale_fill_viridis_d(option = "C", end = 0.9, name = "Month") +
  theme_ridges(font_family = "Roboto Condensed", font_size = 14) +
  theme(legend.position = "none") +
  geom_hline(data = . %>% tail(1), 
             aes(yintercept = vsum), linetype = 2) +
  labs(x="year",y="New Home Sales (1000s NSA)",
       title="U.S. New Home Sales",
       subtitle="Year-to-date total through June",
       caption="@lenkiefer Source: U.S. Census Bureau and Department of Housing and Urban Development\ndotted line at 2018 value") + 
  guides(colour = guide_legend(reverse=FALSE))


# each month col for each year >>> move


```














