---
title: "Untitled"
author: "RN7"
date: "May 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

https://www.kaggle.com/jonathanbouchet/historical-look-at-the-u-s-murders-type/notebook

- murders by type: breakdown by year and State of the murder’s count by type: cutting, firearms, unarmed and other
- firearm murder: details of the murder’s count by firearm: shotgun, handgun, riffle and other

- The data have the State information for only Year after 2004 so I split the data at this date for the by-State study.

- Other than that, we can still aggregate by Year and Type over the main dataframe.

- To normalize the count by State, I will use the census data 2015 from another KAggle dataset: US Census Demographic Data

- For the details by firearm, there is a small correction in the label, i.e some have a s at the end. For example: handgun and handguns which adds another category. So i just removed the s to unify the labels.


```{r cars}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(ggthemes)
library(RColorBrewer)
library(cowplot)
```

read in files

```{r }

all_murders <- read.csv("../murders_analysis_usa/murders by type.csv", sep = ",", stringsAsFactors = FALSE)

# by state information ONLY FROM 2004 >>> split data from 2004 onwards 

df_1 <- all_murders %>% filter(Year >= 2004)
df_2 <- all_murders %>% filter(Year < 2004)

firearms <- read.csv("../murders_analysis_usa/Firearm Murders.csv", sep = ",", stringsAsFactors = FALSE)

firearms %>% glimpse()

sapply(firearms$Type, function(x) gsub('s','',x))

firearms$Type %>% unique()

stringr::str_replace_all(firearms$Type, "s$", " ") %>% trimws() %>% unique()

firearms %>% mutate(Type = stringr::str_replace_all(Type, "s$", "")) %>% select(Type) %>% unique()

firearms <- firearms %>% mutate(Type = stringr::str_replace_all(Type, "s$", ""))

glimpse(firearms)

```


TIMELINE of Murders Count

```{r}
# base plot by Type

basePlot <- all_murders %>% 
  group_by(Year, Type) %>% 
  summarize(count = n(),
            sum = sum(By.Weapon)) %>% 
  ggplot(aes(x = Year, y = sum, group = Type)) +
  geom_line(aes(color = Type, group = Type), size = 1.5, alpha = 0.75) +
  geom_point(size = 1.5) +
  scale_color_brewer(name = "", palette = "Set1") +
  theme_fivethirtyeight() +
  labs(title='Murder\'s count by type',subtitle='refs [1],[2],[3],[4]:https://www.thoughtco.com/us-gun-control-timeline-3963620') +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))

# total count plot

totCount <- all_murders %>% 
  group_by(Year) %>% 
  summarize(tot = sum(By.Weapon)) %>% 
  ggplot(aes(Year, tot)) +
  geom_line(size = 0.5, alpha = 0.5) +
  geom_point(size = 0.5) +
  theme_fivethirtyeight() +
  theme(axis.title = element_blank(), 
        panel.grid.major.y = element_blank(), 
        axis.text = element_text(size = 6)) +
  labs(subtitle = "Total Count") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5),
                     limits = c(0, 25000)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))


```


add in extra annotations! 


```{r fig.width=12, fig.height=8, fig.align='center'}

basePlot +
  geom_curve(aes(x = 1992, xend = 1993, y = 13000, yend = 15900), 
             curvature = 0.05, arrow = arrow(length = unit(0.02, "npc")), color = "grey30", size = 0.25) +
  annotate("text", x = 1992, y = 12000, label = "The Brady Law, \nAssault Weapon Ban[1]", 
           size = 4, color = "black") +
  geom_curve(aes(x = 2000, xend = 1999, y = 11000, yend = 8500), 
             curvature = 0.05, arrow = arrow(length = unit(0.02, "npc")), 
             color = "grey30", size = 0.25) +
  annotate("text", x = 2000, y = 12000, label = "Columbine High School\nMass Shooting [2]",
           size = 4, color = "red") + 
  geom_curve(aes(x = 2006, y = 12000, xend = 2008, yend = 9750), 
             curvature = -0.05, arrow = arrow(length = unit(0.02, "npc")), 
             color='gray30', size=0.25) + 
  annotate("text", x=2006, y = 13000, 
           label = "National Instant Criminal Background\nCheck Improvement Act[3]", 
           size=4, colour="black") +
  geom_curve(aes(x = 2014.75, y = 8000, xend = 2015, yend = 9400), 
             curvature = 0.05, arrow = arrow(length = unit(0.02, "npc")), 
             color='gray30',size=0.25) +
  annotate("text", x=2015, y = 7000, 
           label = "Fix Gun Checks\nAct of 2015 (H.R. 3411)[4]", 
           size=4, colour="black") + 
  geom_curve(aes(x = 2015, y = 12000, xend = 2015.8, yend = 11000), 
             curvature = 0.05, arrow = arrow(length = unit(0.02, "npc")), 
             color='gray30', size=0.25) +
  annotate("text", x=2014.5, y = 13000, 
           label = "Orlando, Florida\nnightclub mass shooting[5]", 
           size=4, colour="black") + 
  annotation_custom(ggplotGrob(totCount),
                    xmin = 2008, xmax = 2017, ymin = 14000, ymax = 19000)
  # place total count graph in top right corner! 

```

tbh I'd just remove the non-firearms line plots >>> focus is on FIREARMS.
added more breaks on x-axis to better understand where important events occur (either that or put YEAR within the annotation text in parentheses). Mini plot of Total murders in top-corner is nice addition.



facet by state: 

```{r}
df_1 %>% 
  group_by(State, Year, Type) %>% 
  summarize(sum = sum(By.Weapon)) %>% 
  ggplot(aes(Year, sum, group = Type )) +
  geom_line(aes(color = Type, group = Type), size = 1, alpha = 0.75) +
  geom_point(size = 1) +
  scale_color_brewer(name = "", palette = "Set1") +
  theme_fivethirtyeight() +
  facet_wrap(~State, ncol = 6) +
  labs(title = 'Murder\'s count by type and State') +
  theme(axis.text = element_text(size = 8))
  
  
```



Normalized MAP for 2016.

```{r fig.width=12, fig.height=10, fig.align='center'}

census_2015 <- read.csv("../murders_analysis_usa/acs2015_census_tract_data.csv", 
                        sep = ",", stringsAsFactors = FALSE)

us_pop_2015 <- census_2015 %>% group_by(State) %>% summarize(totPop = sum(TotalPop))

us_pop_2015 %>% glimpse()

states_map <- map_data("state")

glimpse(states_map)

states_name <- data.frame("state_name" = tolower(state.name),
                          "state_abb" = state.abb)

y2016_df <- all_murders %>% 
  filter(Year == 2016) %>% 
  group_by(State) %>% 
  summarise(sum = sum(By.Weapon))

res <- y2016_df %>% left_join(us_pop_2015, by = "State")
glimpse(res)

res <- res %>% 
  mutate(count_normalized = 100 * (sum / totPop),
         region = tolower(State)) %>% 
  select(-State)
  

all_states_2016 <- left_join(
  states_map, 
  res,
  by = "region")

glimpse(res)
glimpse(states_map)

glimpse(all_states_2016)

# alabama and florida missing for some weird reason

g <- all_states_2016 %>% 
  ggplot(aes(map_id = region)) +
  geom_map(aes(fill = count_normalized), 
           map = states_map, 
           color = "black", size = 0.2) +
  expand_limits(x = states_map$long, y = states_map$lat) + # IMPORTANT for map to appear within grob
  scale_fill_gradientn(name = "", colors = viridis::viridis(10), na.value = "white") +
  theme_fivethirtyeight() +
  theme(panel.grid.major = element_blank(),
        axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        legend.position = c(0.9, 0.02),
        legend.text = element_text(size = 7)) + 
  labs(title = 'Percentage of the murder count normalized by State population for Y2016', 
       subtitle = 'State population from census 2015\nbottom left histogram is the raw count distribution') +
  coord_map(projection = "conic", lat0 = 30)


g1 <- g +
  annotate("text", x = -86, y = 27, 
           label = "Alabama:\nno entry for 2016", 
           size = 4, color = "black") +
  annotate("segment", 
           x = -86, xend = -87, 
           y = 28, yend = 32, 
           color = "black") +
  annotate("text", x = -75, y = 27, 
           label = "Florida: no entry\nin the dataset", 
           size = 4, color = "black")

# set histogram inside map
gg1 <- ggdraw(g1)  # cowplot::ggdraw()

hist <- y2016_df %>% 
  ggplot(aes(sum)) +
  geom_histogram(bins = 20) +
  geom_vline(xintercept = median(y2016_df$sum), 
             color = "orange", size = 1.5) +
  annotate("text", x = 800, y = 10, 
           label = "median value\nover all States", 
           size = 4, color = "orange") +
  theme_fivethirtyeight() +
  theme(axis.text = element_text(size = 8)) +
  labs(subtitle = "un-normalized count distribution")


gg1 + 
  draw_grob(ggplotGrob(hist), 
            x = 0.01, y = 0.14, 
            width = 0.3, height = 0.175)



```


# timeline + breakdown of murders by firearm

```{r}

firearms %>% 
  group_by(Year, Type) %>% 
  summarize(count = n(), 
            sum = sum(By.Type)) %>% 
  ggplot(aes(Year, sum, group = Type)) +
  geom_line(aes(color = Type, group = Type), 
            size = 2, alpha = 0.5) +
  geom_point(size = 2) +
  theme_economist() +
  scale_color_brewer("", palette = "Set1") +
  labs(title = "Murder\'s count by type", 
       subtitle = "all states") +
  facet_grid(~Type, scales = "free_y") +
  geom_smooth(method = "loess", color = "darkred", size = 0.25, alpha = 0.15)

# facet_wrap not work in dev ggplot2 so used facet_grid.......

```



































































