---
title: "Untitled"
author: "RN7"
date: "September 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)

dat <- tibble::tribble(
                                     ~municipality,                                                                                            ~sex_and_collection_number,
                                        "Alegrete", "F: TR917, 939, 959, 963, JR322–325, 333, 334, 399, 400, 413, 416/M: TR960, 964, JR320, 321, 323, 397, 398, 401, 412",
                                  "Alegrete South",                                                                               "F: TR941, 942, 945/M: TR940, 943, 944",
                                        "Candiota",                    "F: TR 177–180, J180, 181, 183, 207, 208, 210, 211, 212, 214–216, 218, 219, 222/M: J209, 213, 217",
                                           "Butiá",                                                             "F: TR289, 594, J131, 170, 199, 201, 204, 205, JR591–595",
                                      "Candelária",                                                                                                        "M: MNHNA1885",
                                "Cachoeira do Sul",                                                                                                        "F: TR921–924"
                          )

glimpse(dat)
dat

```

Place values of sex var into own column >>> split + unnesting, separate on existing delimiter.


```{r}
dat %>% 
  unnest(sex_and_collection_number = strsplit(sex_and_collection_number, "\\/")) %>% 
  separate(sex_and_collection_number, into = c("sex", "specimen"), sep = ": ") -> dat
```

REGEX:

use lookbehind + backreference >>> insert new delimiter in replacement argument before ALL isntances of uppercase characters at word boundary (colelction acronyms in specimen)

```{r}
dat %>% 
  mutate(specimen = str_replace_all(specimen, "((?<!^)\\b[A-Z])", ";\\1")) %>% 
  unnest(specimen = strsplit(specimen, ";")) %>% 
  mutate(specimen = str_squish(specimen)) -> dat
```



collection acronyms into new var based on ; separator

```{r}
dat %>% 
  mutate(museumAbbr = str_extract(specimen, "^[A-Z]+")) %>% 
  mutate(specimen = str_remove(specimen, "^[A-Z]+") %>% str_squish()) %>% 
  mutate(specimen = str_remove(specimen, ",$")) %>% 
  unnest(specimen = strsplit(specimen, ",")) %>% 
  mutate(specimen = str_squish(specimen)) -> dat
```

finally deal with specimens with consecutive ID numbers not explicitly written
- EXPAND the series >>> unnest

```{r}
expand_series <- function(ser_vector, separator) {
  words <- str_split(ser_vector, paste(separator))
  
  if (length(words[[1]]) == 1) {
    return(ser_vector)
  } else {
    start_str <- pluck(words, 1, 1)
  }
  end_str <- pluck(words, 1, 2)
  
  paste0(seq(as.integer(start_str), as.integer(end_str)), collapse = ", ")
}

expand_series <- Vectorize(expand_series)

dat %>% 
  mutate(specimen = expand_series(.$specimen, "-")) %>% 
  unnest(specimen = strsplit(specimen, ", ")) %>% 
  mutate(specimen = str_squish(specimen))


```

