---
title: "Untitled"
author: "RN7"
date: "December 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


FROM: strimas.com

## tidy-sf for mapping

```{r}
library(sf)
library(tidyverse)
library(viridis)
library(rvest)

```

Simple Features: opensource standard for representation of real-world objects

>>> features:
- point
- linestring
- polygon
- etc.

Usage of DATA FRAMES vs. S4 objects in sp package

Column == attribute
Row == feature

list-column `sfc` for storing feature geometries >>> classs `sfg` 

```{r}
nc <- st_read(system.file("shape/nc.shp", package = "sf"))

glimpse(nc)

# data from only first two counties in NC
nc <- nc[1:2, ]
nc <- nc %>% slice(1:2)

# convert to SpatialPolygonsDataFrame
nc_sp <- as(nc, "Spatial")

class(nc)
glimpse(nc)
as_tibble(nc)

# compare to storage methods in sp:

class(nc_sp)
str(nc_sp)

```

Access geometry list-clumn >>> class `sfc` with `st_geometry()`.

```{r}
(nc_geom <- st_geometry(nc))

st_geometry(nc) %>% class()

attributes(nc_geom)

```

Individual simple features == `sfg` objects >>> classes of specific type of feature

```{r}
nc_geom[[1]] %>% class()
# XY and MULTIPOLYGON
```

## Using ggplot with sf objects

sp == conversion with `fortify()` before usage with ggplot2

sf == use `geom_sf()`

```{r}
nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

nc %>% 
  ggplot() +
  geom_sf(aes(fill = AREA)) + 
  scale_fill_viridis("Area") +
  ggtitle("Area of counties in NC") +
  theme_bw()

```

Use `coord_sf()` to plot features in different project

Ex. Albers equal area projection.

```{r}
nc %>% 
  ggplot() +
  geom_sf(aes(fill = AREA)) +
  scale_fill_viridis("Area") +
  coord_sf(crs = st_crs(102003)) +
  ggtitle("Area of countines in NC (Albers projection)") +
  theme_bw()

```

## Using dplyr for data munging

```{r}
nc %>% 
  mutate(area_km2 = AREA * 10000) %>% 
  select(name = NAME, area_km2) %>% 
  filter(area_km2 > 2000) %>% 
  arrange(desc(area_km2)) %>% 
  slice(1:3)


```

geometry column kept EVEN if NOT specified in `select()`, same with `transmute()`! 

Manipulate geometry column with certain sf functions:

```{r}
nc %>% 
  mutate(area_m2 = st_area(geometry)) %>% 
  select(name = NAME, area_m2, area = AREA) %>% 
  head() %>% 
  as.tibble()
```


## Grouped operations with dplyr

```{r}
nc_groups <- nc %>% mutate(group = sample(LETTERS[1:3], nrow(.), replace = TRUE))

nc_mean_area <- nc_groups %>% 
  group_by(group) %>% 
  summarise(area_mean = mean(AREA))

nc_mean_area %>% 
  ggplot() +
  geom_sf(aes(fill = area_mean)) +
  scale_fill_distiller("Area", palette = "Reds") +
  ggtitle("Mean area by group") +
  theme_bw()

```

Attribute AND geometric data aggregated. Geometries within each group combined + boundaries are dissolved! Usage of `sf_union()` internally.

```{r}
nc_groups %>% 
  select(group, AREA) %>% 
  group_by(group) %>% 
  mutate(area_prop = sum(AREA)) %>% 
  as_tibble()

nc_groups %>% 
  select(group, AREA) %>% 
  group_by(group) %>% 
  filter(mean(AREA) > 0.13) %>% 
  as_tibble()


```


## Joins with dplyr

Combine with county-level population data (Wikipedia)

```{r}
url <- "https://en.wikipedia.org/wiki/List_of_counties_in_North_Carolina"

pop <- url %>% 
  read_html() %>% 
  html_nodes("table.wikitable:nth-child(9)") %>% 
  html_table() %>% 
  `[[`(1) %>% 
  select(County, starts_with("Population")) %>% 
  set_names(c("county", "population")) %>% 
  mutate(county = stringr::str_replace(county, " County", ""),
         population = gsub("(^[0-9]*???)|,", "", population) %>% parse_integer()) 
                                              # or use iconv()

glimpse(pop)

pop %>% 
  arrange(desc(population))

nc %>% 
  transmute(county = as.character(NAME)) %>% 
  inner_join(pop, by = "county") %>% 
  ggplot() +
  geom_sf(aes(fill = population)) +
  scale_fill_viridis("Population", labels = scales::comma) +
  ggtitle("County-level population in NC") +
  theme_bw()


```

for joining SPATIAL objects >>> use `st_join()`! 


## tidyr verbs

Use NC births in counties

```{r}
nc_gathered <- nc %>% 
  select(couunty = NAME, BIR74, BIR79, -geometry) %>% 
  slice(1:3) %>% 
  gather(key = year, value = births, BIR74, BIR79)

nc_gathered

nc_gathered %>% 
  spread(year, births)

```






# Lobster article:


```{r}
library(here)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(hrbrthemes)
library(gganimate)


cols <- c("year", "state", "species", "tons", "pounds", "value")

lobsters <- here("lobsters/") %>% 
  list.files("^lobster", full.names = TRUE) %>% 
  map_df(read_csv, skip = 5, col_names = cols) %>% 
  mutate(
    species = recode(species, 
                     "LOBSTER, AMERICAN" = "American Lobster",
                     "LOBSTER, BANDED SPINY" = "Banded Spiny Lobster",
                     "LOBSTER, CALIFORNIA SPINY" = "California Spiny Lobster", 
                     "LOBSTER, CARIBBEAN SPINY" = "Caribbean Spiny Lobster", 
                     "LOBSTER, SLIPPER" = "Slipper Lobster"),
    state = recode(state, 
                   "Florida East Coast" = "Florida",
                   "Florida West Coast" = "Florida")) %>% 
  group_by(year, state, species) %>% 
  summarize_all(funs(sum)) %>% 
  filter(state != "Hawaii") %>% 
  ungroup()

glimpse(lobsters)

```


Lobster catch by species:

```{r}

lobster_trend <- lobsters %>% 
  group_by(year, species) %>% 
  select(-state) %>% 
  summarize_all(funs(sum))

lobster_trend %>% 
  ggplot(aes(year, tons / 1000)) +
  geom_line(aes(color = species, group = species), size = 1.15) + # add group = species
  scale_color_brewer(NULL, palette = "Set1") +
  scale_y_comma("Catch (millions kg)") +
  scale_x_continuous("Year", breaks = seq(1950, 2010, by = 10)) +
  labs(title = "US commercial catch of lobster by species") +
  theme_bw() +
  theme(legend.position = "bottom")



```

https://kohske.wordpress.com/2010/12/27/faq-geom_line-doesnt-draw-lines/


Catch by state over time.

Top_9 states:

```{r}
top_9_states <- lobsters %>% 
  filter(species == "American Lobster") %>% 
  group_by(state) %>% 
  select(-year, -species) %>% 
  summarize_all(funs(sum)) %>% 
  top_n(9, tons) %>% 
  arrange(desc(tons)) %>% 
  pull(state)      # grab the values in specified column

lobster_state <- lobsters %>% 
  filter(species == "American Lobster",
         state %in% top_9_states) %>% 
  mutate(state = factor(state, levels = top_9_states))

lobster_state %>% 
  ggplot(aes(year, tons / 1000)) +
  geom_line(aes(color = state), size = 1.15) +
  scale_color_brewer(palette = "Set1") +
  scale_y_comma("Catch (millions, kg)") +
  scale_x_continuous("Year", breaks = seq(1950, 2010, by = 10)) +
  labs(title = "US commercial catch of American Lobster") +
  theme_bw() +
  theme(legend.position = "bottom")
  

```

MAINE! 


## Animated map

```{r}
# state boundaries
states <- ne_states(iso_a2 = c("US", "CA"), returnclass = "sf") %>% 
  filter(!postal %in% c("AK", "HI")) %>%  # filter out alaska and hawaii
  select(country = iso_a2, state = name, state_code = postal, geometry)

glimpse(states)
str(states)

# join with lobster data
lobster_sf <- lobsters %>% 
  filter(species == "American Lobster") %>% 
  complete(state, year, fill = list(tons = 0, pounds = 0, value = 0)) %>% # fill in NAs
  select(-species) %>% 
  inner_join(states, ., by = "state")

glimpse(lobster_sf)

# log scaling
rng <- ceiling(range(log10(lobster_sf$tons + 1)))
brk <- rng[1]:rng[2]
lbl <- c(0, bquote(10^1), bquote(10^2), bquote(10^3), bquote(10^4), bquote(10^5))

proj <- paste0("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96",
               "+x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")

g <- ggplot() +
  geom_sf(data = states, fill = "grey80", color = "grey60") +
  geom_sf(data = lobster_sf,
          aes(fill = log10(tons + 1), frame = year)) +
  scale_fill_viridis("Annual commercial lobster catch (metric tons)",
                     breaks = brk, labels = lbl, 
                     guide = guide_colorbar(title.position = "top",
                                            title.hjust = 0.5)) +
  coord_sf(crs = proj,
           xlim = c(553469.1, 2259794),
           ylim = c(-259859.5, 1406546.0)) +
  labs(title = NULL, caption = "NOAA") +
  theme_bw() +
  theme(legend.position = "bottom", 
        legend.key.width = unit(0.15, "npc"),
        plot.title = element_text(hjust = 0.5))

gganimate(g, interval = 0.2)


```




## Birds


```{r}
library(tidyverse)
library(rvest)
library(stringr)
library(sf)
library(leaflet)
library(leaflet.extras)
library(USAboundaries)
library(rnaturalearth)
library(viridis)
library(here)
```


NA regional statistics

```{r}
ebird_regions <- data_frame(
  region_code = c("CA", "US", "MX"),
  region_name = c("Canada", "United States", "Mexico"), 
  region_level = "country")

country_data <- function(country) {
  base_url <- "http://ebird.org/ebird/country/%s?yr=all"
  
  page <- sprintf(base_url, country) %>% 
    read_html()
  
  counts <- page %>% 
    html_nodes(".hs-section-count") %>% 
    html_text() %>% 
    str_extract("[0-9,]+") %>% 
    parse_number()
  
  data_frame(n_species = counts[1], 
             n_checklists = counts[2])
  
}

ebird_regions <- ebird_regions %>% 
  mutate(country_df = map(region_code, country_data)) %>% 
  unnest()

```

## Extraction function

Extracts all sub-region data within region
>>> use `Sys.sleep()` function to pause between each page load


```{r}
extract_subregion_data <- function(region_code, region_level, sleep = 5) {
  Sys.sleep(sleep)
  
  base_url <- "http://ebird.org/ebird/%s/%s/regions?yr=all"
  
  page <- sprintf(base_url, region_level, region_code) %>% 
    read_html()
  
  region_stats <- html_table(page)
  # skip this region if no data at sub-region
  if (length(region_stats) == 0) {
    return(data_frame())
  }
  
  region_stats <- region_stats[[1]] %>% 
    set_names(c("rank", "region_name", "n_species", "n_checklists"))
  
  # region codes
  region_urls <- page %>% 
    html_nodes("td a") %>% 
    html_attr("href")
  
  region_codes <- region_urls %>% 
    str_extract("[-A-Z0-9]+\\?") %>% 
    str_replace("\\?", "")
  
  region_levels <- region_urls %>% 
    str_extract("ebird/[a-z0-9]+") %>% 
    str_replace("ebird/", "")
  
  mutate(region_stats, 
         region_code = region_codes, 
         region_level = region_levels) %>% 
    select(region_code, region_name, region_level,
           n_species, n_checklists) %>% 
    as_tibble()
}

# state data
state_df <- map_df(ebird_regions$region_code, extract_subregion_data,
                   region_level = "country")

# country data
county_df <- map_df(state_df$region_code, extract_subregion_data, 
                    region_level = "subnational1")

# combine all

ebird_regions <- ebird_regions %>% bind_rows(state_df, county_df)

glimpse(ebird_regions)

getwd()
write.csv(ebird_regions, "ebird_regions.csv")

```

## Top hotspots 


```{r}
extract_top_hotspot <- function(region_code, region_level, sleep = 5) {
  
  Sys.sleep(sleep)
  
  base_url <- "http://ebird.org/ebird/%s/%s/hotspots?yr=all"
  
  th <- sprintf(base_url, region_level, region_code) %>% 
    read_html() %>% 
    html_node("td a") %>% 
    html_attr("href") %>% 
    str_extract("L[0-9]+")
  
  if (!is.character(th) || length(th) != 1) {
    return(NA_character_)
  } else {
    return(th)
  }
  
}

ebird_regions <- ebird_regions %>% 
  mutate(top_hotspot = map2_chr(region_code, region_level, extract_top_hotspot))

str(county_df)

```









