---
title: "Untitled"
author: "RN7"
date: "May 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## https://walkerke.github.io/2017/05/tigris-metros/

```{r}
library(tigris)
library(sf)
library(tidyverse)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

Grab tracts for Portland, Oregon metro area.

Combine with `rbind_tigris()` function!

```{r}
orwa <- rbind_tigris(
  tracts("OR", cb = TRUE),
  tracts("WA", cb = TRUE)
)


ggplot(orwa) +
  geom_sf()

```

Sbuset tracts by locating boundary of Portland metro area >>> available in `core_based_statistical_areas()` function. Find "Portland-Vancouver" string in dataset.

```{r}
cb <- core_based_statistical_areas(cb = TRUE)

glimpse(cb)

pdx <- cb %>% filter(str_detect(NAME, "Portland-Vancouver"))

lax <- cb %>% filter(str_detect(NAME, "Los Angeles-Long Beach-Anaheim, CA"))

ggplot(pdx) + geom_sf()
ggplot(lax) + geom_sf()

```

Spatial subsetting by indexing data 

```{r}
p1 <- orwa[pdx, ]

ggplot() +
  geom_sf(data = p1) +
  geom_sf(data = pdx, fill = NA, color = "darkred")

```

Returned tracts within metro area but ALSO areas outside that touch metro area boundary. 

Use `st_within()` to ID tracts WITHIN boundary of the metro area.

```{r}
w1 <- orwa %>% st_within(pdx)

print(length(w1))
print(w1[1:5])
```

`w1` is LIST with same length as `orwa`. Values == 0 for NOT within metro or 1 == WITHIN  metro area.

Use `map_lgl()` to convert to logical vector and subset according to vector.

```{r}
w2 <- 
  map_lgl(w1, function(x) {
    
    if (length(x) == 1) {
      return(TRUE)
    } else {
      return(FALSE)
    }
  }) 

# w2 == length of w1 and orwa

p2 <- orwa[w2, ]

ggplot() +
  geom_sf(data = p2) +
  geom_sf(data = pdx, fill = NA, color = "red")


```

Custom function for OTHER metro areas.

1. Metro area as INPUT. Detect states wherein metro are located.
2. Retrieve tracts from states.
3. IDs tracts located WITHIN metro area boundary.


```{r}
metro_tracts <- function(metro_name) {
  
  # ID which states intersect metro area with `states()` function
  st <- states(cb = TRUE)
  
  cb <- core_based_statistical_areas(cb = TRUE)
  
  metro <- cb %>% filter(str_detect(NAME, metro_name))
  
  stcodes <- st[metro,]$STATEFP
  
  
  # fetch tracts, use `rbind_tigris()` for more than one state
  if (length(stcodes) > 1) {
    
    tr <- rbind_tigris(
      map(stcodes, function(x) {
        tracts(x, cb = TRUE)
      })
    )
  } else {
    tr <- tracts(x, cb = TRUE)
  }
  
  
  # find tracts within metro area
  within <- st_within(tr, metro)
  
  within_lgl <- map_lgl(within, function(x) {
    if(length(x) == 1) {
      return(TRUE)
    } else {
      return(FALSE)
    }
  })

  # output subset return
  output <- tr[within_lgl, ]
  
  return(output)
}

# must be valid name within "cb"
la <- metro_tracts("los angeles")   # object 'x' NOT FOUND
la <- metro_tracts("Los Angeles")

chi <- metro_tracts("Chicago")
sf <- metro_tracts("San Francisco")

```


```{r}
metro_tracts <- function(metro_name) {
  
  # First, identify which states intersect the metro area using the
  # `states` function in tigris
  st <- states(cb = TRUE)
  cb <- core_based_statistical_areas(cb = TRUE)
  metro <- filter(cb, grepl(metro_name, NAME))
  
  stcodes <- st[metro,]$STATEFP
  
  # Then, fetch the tracts, using rbind_tigris if there is more
  # than one state
  if (length(stcodes) > 1) {
    tr <- rbind_tigris(
      map(stcodes, function(x) {
        tracts(x, cb = TRUE)
      })
    )
  } else {
    tr <- tracts(x, cb = TRUE)
  }
  
  # Now, find out which tracts are within the metro area
  within <- st_within(tr, metro)
  
  within_lgl <- map_lgl(within, function(x) {
    if (length(x) == 1) {
      return(TRUE)
    } else {
      return(FALSE)
    }
  })
  
  # Finally, subset and return the output
  output <- tr[within_lgl,]
  
  return(output)
  
}
```

























































