---
title: "Untitled"
author: "RN7"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://www.cultureofinsight.com/blog/2018/05/02/2018-04-08-multivariate-dot-density-maps-in-r-with-sf-ggplot2/

UK Election 2017 >>> LONDON


```{r}
library(ggplot2)
library(dplyr)
library(sf)
library(readr)
extrafont::loadfonts("win")


```

load data

```{r}
# election results
ge_data <- read_csv("http://researchbriefings.files.parliament.uk/documents/CBP-7979/HoC-GE2017-constituency-results.csv") %>% 
  filter(region_name == "London") %>% 
  select(ons_id, constituency_name, first_party, Con = con, Lab = lab, LD = ld, UKIP = ukip, Green = green)

# shapefiles >>> filter only LONDON
uk <- st_read("../uk_london_shapefiles/uk_650_wpc_2017_full_res_v1.8.shp", 
              stringsAsFactors = FALSE, 
              quiet = TRUE) %>% 
  st_transform(4326) %>% 
  filter(REGN == "London") %>% 
  select(ons_id = PCONCODE)

# merge dataframes

sf_data <- uk %>% 
  right_join(ge_data, by = "ons_id")

head(sf_data)

```

random_round
dataframe for dots (100 votes) for each party

```{r}
random_round <- function(x) {
  
  v <- as.integer(x)
  
  r <- x - v
  
  test <- runif(length(r), 0.0, 1.0)
  
  add <- rep(as.integer(0), length(r))
  
  add[r > test] <- as.integer(1)
  
  value <- v + add
  
  ifelse(is.na(value) | value < 0, 0, value)
  
  return(value)
  
}

num_dots <- as.data.frame(sf_data) %>% 
  select(Con:Green) %>% 
  mutate_all(funs(. / 100)) %>% 
  mutate_all(random_round)

nu <- as.data.frame(sf_data) %>% 
  select(Con:Green) %>% 
  mutate_all(funs(. / 100))

glimpse(num_dots)


# dataframe with coordinates for each point + party
library(purrr)
sf_dots <- map_df(
  names(num_dots),
  ~ st_sample(sf_data, size = num_dots[, .x], type = "random") %>% # generate points for each polygon
    st_cast("POINT") %>% # cast as point data
    st_coordinates() %>% # pull out coordinates into matrix
    as_tibble() %>% 
    setNames(c("lon", "lat")) %>% # set col names to longitude, latitude
    mutate(Party = .x) 
  ) %>% 
  slice(sample(1:n())) # bind rows + randomise order to avoid bias in plotting order


head(sf_dots)


```

color palette and plotting :: need larger image size and adjust text/legend accordingly

```{r fig.height=20, fig.width=20}
pal <- c("Con" = "#0087DC", "Lab" = "#DC241F", "LD" = "#FCBB30", 
         "UKIP" = "#70147A", "Green" = "#78B943")

p <- ggplot() +
  geom_sf(data = sf_data, fill = "transparent", color = "white") +
  geom_point(data = sf_dots, aes(lon, lat, color = Party)) +
  scale_color_manual(values = pal) +
  coord_sf(crs = 4326, datum = NA) +
  theme_void(base_family = "Iosevka", base_size = 48) +
  labs(x = NULL, y = NULL,
       title = "UK General Election 2017\n",
       subtitle = "London Constituencies\n1 dot = 100 votes",
       caption = "Map by Culture of Insight @PaulCampbell91 | Data Sources: House of Commons Library, Alasdair Rae") +
  guides(color = guide_legend(override.aes = list(size = 18))) +
  theme(legend.position = c(0.82, 1.03), 
        legend.direction = "horizontal",
        plot.background = element_rect(fill = "#212121", color = NA),
        panel.background = element_rect(fill = "#212121", color = NA),
        legend.background = element_rect(fill = "#212121", color = NA),
        legend.key = element_rect(fill = "#212121", color = NA),
        plot.margin = margin(1, 1, 1, 1, "cm"),
        text = element_text(color = "white"),
        title = element_text(color = "white"),
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(size = 32)
        )
 
p


```

Compare against choropleht map 

```{r}
ggplot() +
  geom_sf(data = sf_data, aes(fill = first_party), 
          color = "white") +
  scale_fill_manual(values = pal, name = "Seat Winner") +
  coord_sf(crs = 4326, datum = NA) +
  theme_void() +
  theme(legend.position = c(0.8, 0.9), 
        legend.direction = "horizontal")
```


What do we think is the most insightful map? Luckily we don’t have to choose one or the other, can use both! No one map will be able to give the you all the answers so I find that it’s best to combine techniques for maximum insight. The choropleth gives us a clear indication as to who won where, and the dot-density looks under the hood and gives us an idea of the count and diversity of votes within each constituency.







